
\chapter*{Vita}
\addcontentsline{toc}{chapter}{Vita}

\textbf{James Chuang}

\begin{description}[align=right, labelwidth=4cm, noitemsep, leftmargin=!]
    \item [year of birth:] 1991
    \item [contact address:] 77 Avenue Louis Pasteur\\
                             Room 239\\
                             Boston, MA 02115
\end{description}

\noindent\hrulefill

\section*{Education}
\begin{description}[align=right, labelwidth=1.5cm, noitemsep, leftmargin=!]
    \item [2018] MSc, Biomedical Engineering, Boston University
    \item [2013] BSc, Biomedical Engineering, Johns Hopkins University
\end{description}


\section*{Publications}
\begingroup
\singlespacing
\begin{description}[align=right, labelwidth=1.5cm, itemsep=1em, leftmargin=!]
    \item [2018] Doris SM*, \textbf{Chuang J}*, Viktorovskaya O, Murawska M, Spatt D, Churchman LS, Winston F (2018). Spt6 is required for the fidelity of promoter selection. \textbf{Molecular Cell}, doi:\href{https://doi.org/10.1016/j.molcel.2018.09.005}{10.1016/j.molcel.2018.09.005}
    \item [2018] \textbf{Chuang J}, Boeke JD, Mitchell LA (2018). Coupling Yeast Golden Gate and VEGAS for Efficient Assembly of the Violacein Pathway in \textit{Saccharomyces cerevisiae}. \textbf{Synthetic Metabolic Pathways}, doi:\href{https://doi.org/10.1007/978-1-4939-7295-1_14}{10.1007/978-1-4939-7295-1\_14}
    \item [2017] Aquino P, Honda B, Suma Jaini, Lyubetskaya A, Hosur K, Chiu JG, Ekladious I, Hu D, Jin L, Sayeg MK, Stettner AI, Wang J, Wong BG, Wong WS, Alexander SL, Ba C, Bensussen SI, Chou K, \textbf{Chuang J}, Gastler DE, Grasso DJ, Greifenberger JS, Guo C, Hawes AK, Israni DV, Jain SR, Kim J, Lei J, Li H, Li D, Li Q, Mancuso CP, Mao N, Masud SF, Meisel CL, Mi J, Nykyforchyn CS, Park M, Peterson HM, Ramirez AK, Reynolds DS, Rim NG, Saffie JC, Su H, Su WR, Su Y, Sun M, Thommes MM, Tu T, Varongchayakul N, Wagner TE, Weinberg BH, Yang R, Yaroslavsky A, Yoon C, Zhao Y, Zollinger AJ, Stringer AM, Foster JW, Wade J, Raman S, Broude N, Wong WW, Galagan JE (2017). Coordinated regulation of acid resistance in \textit{Escherichia coli}. \textbf{BMC Systems Biology}, doi:\href{https://doi.org/10.1186/s12918-016-0376-y}{10.1186/s12918-016-0376-y}
    \item [2015] Mitchell, LA*, \textbf{Chuang J}*, Agmon N, Khunsriraksakul C, Phillips NA, Cai Y, Truong DM, Veerakumar A, Wang Y, Mayorga M, Blomquist P, Sadda P, Trueheart J, Boeke JD (2015). Versatile genetic assembly system (VEGAS) to assemble pathways for expression in \textit{S. cerevisiae}. \textbf{Nucleic Acids Research}, doi:\href{https://doi.org/10.1093/nar/gkv466}{10.1093/nar/gkv466}
    \item [2015] Agmon N, Mitchell LA, Cai Y, Ikushima S, \textbf{Chuang J}, Zheng A, Choi W, Martin JA, Caravelli K, Stracquadanio G, Boeke JD (2015). Yeast Golden Gate (yGG) for the Efficient Assembly of \textit{S. cerevisiae} Transcription Units. \textbf{ACS Synthetic Biology}, doi:\href{https://doi.org/10.1021/sb500372z}{10.1021/sb500372z}
    \item [2013] Mitchell LA, Cai Y, Taylor M, Noronha AM, \textbf{Chuang J}, Dai L, Boeke JD (2013). Multichange isothermal mutagenesis: a new strategy for multiple site-directed mutations in plasmid DNA. \textbf{ACS Synthetic Biology}, \\doi:\href{https://doi.org/10.1021/sb300131w}{10.1021/sb300131w}
\end{description}
\endgroup
